---
title: "R Notebook"
output: html_notebook
---

# Mario Kart 64 World Records
# Exploratory Analysis

# Questions
+ How did the world records develop over time?
+ Which track is the fastest?
+ For which track did the world record improve the most?
+ For how many tracks have shortcuts been discovered?
+ When were shortcuts discovered?
+ On which track does the shortcut save the most time?
+ Which is the longest standing world record?
+ Who is the player with the most world records?
+ Who are recent players?

# Data
Tidytuesday data set can be found [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-05-25/readme.md). 

# Methodology
Use R tidyverse to explore and make vizualizations for tidytuesday data

# Results

# Limitations and Next Steps
Making interactive plots would be the next step.

```{r setup, echo=FALSE,include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(include = FALSE,echo = FALSE,warning = FALSE)

records <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/records.csv')
drivers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/drivers.csv')

head(records)
head(drivers)
```

```{r records-over-time}
p1 <- records %>% 
  left_join(drivers,by="player") %>% 
  group_by(year) %>% 
  summarise(records = mean(records,na.rm=TRUE)) %>% 
  ggplot(aes(year,records)) +
  geom_line(color="darkgreen") + 
  ggthemes::theme_hc() +
  labs(y = "Average Records",
       title = "Average Records Over Time")

# Seems the records went down after year 2000
```

```{r fastest-track}
records %>% 
  group_by(track) %>% 
  summarise(time = median(time)) %>% 
  ggplot(aes(reorder(track,time),time)) +
  geom_col(fill="darkgreen") +
  coord_flip() +
  ggthemes::theme_hc() +
  labs(y = "Median Time(s)",
       x = "",
       title = "Median Time per Track")

p2 <- records %>% 
  group_by(track) %>% 
  summarise(time = mean(time)) %>% 
  ggplot(aes(reorder(track,time),time)) +
  geom_col(fill="darkgreen") +
  coord_flip() +
  ggthemes::theme_hc() +
  labs(y = "Average Time(seconds)",
       x = "",
       title = "Average Time per Track")
```

```{r shortcut-discovered}
# 12 tracks have shortcuts
records %>% 
  mutate(shortcut = as.factor(shortcut),
         track = as.factor(track)) %>% 
  group_by(track) %>% 
  summarise(shortcut = sum(shortcut == "Yes")) %>% 
  arrange(shortcut)
```

```{r save-most-time}
p3 <- records %>% 
  ggplot(aes(reorder(track,time),time)) +
  geom_col(aes(fill=shortcut),position = "dodge") +
  coord_flip() +
  facet_wrap(~shortcut) +
  theme(legend.position = "none") +
  labs(x = "",
       title = "Time For Each Track by Shortcut")

records %>% 
  mutate(track = as.factor(track),
         shortcut= as.factor(shortcut)) %>% 
  filter(track != "Banshee Boardwalk",
         track != "Bowser's Castle",
         track!="Koopa Troopa Beach",
         track != "Moo Moo Farm") %>%  # Remove tracks with no shortcuts
  group_by(track) %>% 
  summarise(total_time_no = sum(time[shortcut=="No"]),
            total_time_yes = sum(time[shortcut=="Yes"]),
            dtime = sum(time[shortcut=="No"]-time),
            mean_time_no = mean(time[shortcut=="No"]),
            mean_time_yes = mean(time[shortcut=="Yes"]),
            diff_mean_time = mean_time_no - mean_time_yes) %>% 
  ggplot(aes(reorder(track,diff_mean_time),diff_mean_time)) +
  geom_col() +
  coord_flip() +
  labs(x = "",
       y = "Difference in Mean Time",
       title = "Mean Difference by Shortcut For Each Track") +
  ggthemes::theme_hc()
  
```

```{r longest-record}
p4 <- records %>% 
  left_join(drivers,by="player") %>% 
  group_by(track,player) %>% 
  summarise(max = max(record_duration)) %>% 
  arrange(desc(max)) %>% 
  head() %>% 
  ggplot(aes(track,max,fill=player)) +
  geom_col(position = "dodge") + 
  coord_flip() + 
  scale_fill_manual(values= wesanderson::wes_palette(n=4, name="GrandBudapest1")) +
  ggthemes::theme_hc() +
  labs(x = "",
       y = "Max Record Duration",
       title = "Max Record Duration by Track and Player")
  
```

```{r players-with-the-most-world-records}
p5 <- drivers %>% 
  group_by(player) %>%
  summarise(record_count = sum(records,na.rm = TRUE)) %>% 
  arrange(desc(record_count)) %>% 
  head(5) %>% 
  ggplot(aes(reorder(player,record_count),record_count,fill=player)) +
  geom_col() +
  coord_flip() +
  labs(x = "",
       y = "Record Count",
       title = "Top 5 players with the most world records") +
  ggthemes::theme_hc() +
  scale_fill_manual(values= wesanderson::wes_palette(n=5, name="BottleRocket1")) +
  theme(legend.position = "none")
```

```{r arrange-plots, fig.height=5}
ggpubr::ggarrange(p1,p2,p3,p4,p5,ncol=2,nrow = 3)
```

